Splitting of the homology of the punctured mapping class group

### 2018-11-05 Actiongg.tex
First draft ready, sent to cfb,palmer,kranhold,kiefner,soulie for comments

\subsection{Bidigrees of the form $(m,m)$}
In this subsection we construct a $\gg-$equivariant isomorphism 
\[
\psi_m\colon \Sym_m(\H))\to H_m(\cms).
\]
% the former group is isomorphic to $\Sym_m(H^1(\S))$ and the second to $H^m(\cms)$,
% both by Poincaré-Lefschetz duality.

We already know that these $\Z_2-$vector spaces have the same dimension:
for the dimension of $H_m(\cms)$ we have to count the monomials of degree $m$ and weight $m$
in the right-hand side of equation \ref{eq:isovectorspaces}; a monomial whose
weight is \emph{equal} and not bigger than its degree cannot contain factors of the form
$Q^i\epsilon$, i.e. it must lie in the factor $\Sym_{\bullet}(\H)$, and if the degree
is $m$ it must lie in $\Sym_m(\H)$.
% 
% The rough idea in constructing the map $\psi_m$ is the following: $\Sym_m(\H)$ is generated
% by elements of the form $[c_1]\cdot\ldots\cdot[c_m]$, where $c_1,\dots,c_m\colon\Sone\to\mrS$ are simple
% closed curves; we consider the product map 
% \[
% c_1\times\dots\times c_m\colon \pa{\Sone}^{\times m}\to \SP^m\pa{\mrS}
% \]
% which carries the fundamental class of the $m-$fold torus to some class in
% $H_m(\SP^m\pa{\mrS})$; we then lifts this class, in a nice way, to a class
% in $H_m(\cms)$.

@@

\begin{defn}
\label{defn:cCm}
Let $\cC^m$ be the the (discrete) set of isotopy classes of $m-$tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
such that:
\begin{itemize}
%  \item each $c_r\colon\Sone\to\mrS$ is smooth and transverse to each $\U_i$ and $\V_i$;
 \item the curves $\set{c_i}_{i\leq m}$ intersect each other only inside $\D$;
 \item each curve $c_i$ is transverse to $\partial\D$.
\end{itemize}
Two such $m-$tuples are isotopic if there is an ambient isotopy of $\S$ relative to $\partial\S\cup\D$
transforming one $m-$tuple into the other. In particular $\cC^m$ is more than countable.
% For each isotopy class we
% suppose to have fixed a representative $(c_1,\dots,c_m)$ made of curves that are
% also transverse to all $\U_i$'s and $\V_i$'s. We write
% $[c_1,\dots,c_m]$ for the isotopy class whose preferred representative is
% $(c_1,\dots,c_m)$.

We denote by $\ZcC{m}$ the free $\Z_2-$vector space with basis $\cC^m$.
There is a canonical surjective map $\pr_m\colon\ZcC{m}\to\Sym_m(\H)$ given by
\[
 \pr_m([c_1,\dots,c_m])=[c_1]\cdot\ldots\cdot[c_m],
\]
where $[c_i]\in\H$ is the fundamental class of the curve $c_i$.

The group $\gg$ acts both on $\ZcC{m}$, by acting on its basis $\cC^m$, and
on $\Sym_m(\H)$, symplectically. The map $\pr_m$ is $\gg-$equivariant.
\end{defn}
% Our aim is to define a $\gg-$equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
% factors through $\pr$ to give the desired $\gg-$equivariant map $\psi$; we will then show that
% $\psi$ is an isomorphism.

% The map $\tpsi_m$ that we construct will have two additional properties. The first
% is the following: for any
% $(c_1,\dots,c_m)$ representing a class in $\cC^m$, the homology
% class $\tpsi_m(c_1,\dots,c_m)$ can be represented by a singular cycle in $\cms$ which
% is \emph{supported} on the union $\D\cup c_1\cup \dots\cup c_m$, meaning that
% this singular cycle only hits configurations in $\cms$ whose points actually
% lie in $\D\cup c_1\cup \dots\cup c_m$.
% 
% For the second property we need a definition
\begin{defn}
 \label{defn:cmsD}
 The space $\cmsD$ is the subspace of $\SP^{m}\pa{\mrS}$ of configurations where all points
 having multiplicity $\geq 2$ lie inside $D$. The space $\cmsD$ is an open subset
 of $\SP^m\pa{\mrS}$, which is a $2m-$dimensional manifold, hence $\cmsD$ is also a manifold.

 There is a natural inclusion $\iota_m\colon\cms\subset\cmsD$ and there is a natural map
  $\j_m\colon \ZcC{m}\to H_m(\cmsD)$ defined as follows: for a class in $\cC^m$ represented
  by a $m-$tuple $(c_1,\dots,c_m)$, the composition
  \[
   \begin{CD}
    \pa{\Sone}^{\times m} @>c_1\times\dots\times c_m >> \pa{\mrS}^{\times m} @>>> \SP^m\pa{\mrS}
   \end{CD}
  \]
has image in the subspace $\cmsD$; we define $\j_m(c_1,\dots,c_m)$ as the image
of the fundamental class
of the $m-$fold torus in $H_m(\cmsD)$. The result doesn't change
if we substitute $(c_1,\dots,c_m)$ with another $m-$tuple which is isotopic relative
to $\D\cup \partial\S$.

@@

\begin{lem}
 \label{lem:cms->cmsDinj}
 The inclusion $\iota_m\colon\cms\to\cmsD$ induces an injective map
 \[
  \pa{\iota_m}_*\colon H_m(\cms)\to H_m(\cmsD).
 \]
\end{lem}
\begin{proof}
 It is equivalent to prove that the map $\iota_m^*\colon H^m(\cmsD)\to H^m(\cms)$ is surjective,
 or, using Poincaré-Lefschetz duality, that the map $H_m(\cmsD^c)\to H_m(\cms^c)$
 is surjective: this last map is induced by the map $\cmsD^c\to\cms^c$ collapsing the subspace
 $\cmsD^c\setminus\cms$ to $\infty$.

 Recall from lemma \ref{lem:gensymchain} that a $\Z_2-$basis for $H_m(\cms^c)$
 is given by classes $[p,(\alpha_j),(u_i,v_i)]$ with $p=0$, $\alpha_j=0$ for all $j$,
 and $\sum_{i=1}^g (u_i+v_i)=m$. Here and in the following we abbreviate $[0,(u_i,v_i)]$.

@@

 It is straightforward to check that the map $\phi^{\tup}$ factors through
 $(\cmsD,\infty)$, as a map of pairs; surjectivity of $H_m(\cmsD^c)\to H_m(\cms^c)$ follows.
\end{proof}

@@

We now prove that the map $\j_m$ lifts along $(\iota_m)_*$
to a $\gg-$equivariant map $\tpsi_m$ as in the diagram.
Since $(\iota_m)_*$ is injective by lemma \ref{lem:cms->cmsDinj}, it suffices to prove that $\j_m$
lands in the image of $(\iota_m)_*$, and the last statement doesn't depend on how $\gg$ acts on these groups.

@@

\begin{defn}
\label{defn:variationsCm}
We introduce a few variations of the notion of configuration space.
\begin{itemize} 
 \item The space $C_{1,m}(\S)$ is the subspace of $\mrS\times \cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots,P_m}}$ with $\bar P\neq P_i$ for all $i$; in other words it is
 the space of configurations of $m+1$ points, one of which is \emph{red} (meaning that
 it is distinguishable from the other points), whereas the other are \emph{black} and not distinguishable
 from each other.
%  \item The space $\fmsD$ is the subspace of $\mrS\times\cmsD$ containing all configurations
%  $(q;\set{p_1,\dots,p_m})$ where either $q\in\D$ may coincide with some $p_i$ (which can appear
%  with multipliciy $\geq 1$), or $q\not\in\D$ must be distinct from all $p_i$'s.
%  Again $q$ is the red point
 \item The space $\fmstwoD$ is the subspace of $\mrS\times\cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots, P_m}}$ where either $\bar P\in\D$ may coincide with \emph{exactly}
 one $P_i$, or
 $\bar P\not\in \D$ must be distinct from all $P_i$'s.
 Again $\bar P$ is called the red point.
 %We see that $C_{1,m}\subset\fmstwoD$ is an open subspace.
 \item The space $\cmstwoD$ is the subspace of $\SP^{m+1}\pa{\mrS}$ of configurations where either all $m+1$ points
 are distinct, or there is exactly one point \emph{inside $\D$} with multiplicity $2$ and $m-1$ other points,
 somewhere in $\S$, with multiplicity $1$.
\end{itemize}
 \end{defn}

We the following natural inclusions:
\[
C_{1,m}(\S)\subset\fmstwoD\subset\fmsD\subset\mrS\times\cmsD\subset\mrS\times\SP^m\pa{\mrS};
\]
\[
C_{m+1}(\S)\subset\cmstwoD\subset C_{m+1}^{\D}(\S)\subset\SP^{m+1}\pa{\mrS}.
\]
All these spaces are
manifolds of dimension $2m+2$ and all inclusions are open.

@@

Fix a small neighborhood $\N$ of $\D\cup c_1\cup\dots\cup c_{m+1}$.

@@

There is a natural map $\p\colon \fmstwoD\to \cmstwoD$, which converts the red point
into a black point. This map restricts to %maps $\fmstwoD\to\cmstwoD$ and 
a map $C_{1,m}(\S)\to C_{m+1}(\S)$;
% the latter being even a $m+1-$fold covering map;
we have a commutative diagram
 \begin{equation}\label{eq:cmstwodiagram}
  \begin{CD}
   C_{1,m}(\S) @>\subset >> \fmstwoD %@>\subset >> \fmsD
\\   @V\p VV @V\p VV %@V\p VV
\\   C_{m+1}(\S) @>\subset >> \cmstwoD %@>\subset >> C_{m+1}^D(\S)
   \end{CD}
\end{equation}

\begin{defn}
\label{defn:falsediagonals} 
Let 
\[
\Dmone=\cmstwoD\setminus C_{m+1}(\S)
\]
and similarly
\[
\Donem=\fmstwoD\setminus C_{1,m}(\S).
\]
We notice that $\p$ restricts to a homeomorphism $\Donem\to\Dmone$. Moreover both $\Dmone\subset\cmstwoD$
and $\Donem\subset\fmstwoD$ are properly embedded submanifolds of codimension $2$, and the map $\p$ restricts to
a $2-$ramified covering between their respective normal bundles.
\end{defn}

@@

We first notice that there is a lift $[c_1]\otimes\fc$ to a cycle in $\pa{\fmstwoD}$,
which is defined by declaring the point in $[c_1]\cdot\fc$ that spins around $c_1$
to be red. We then notice that the right vertical map
\[
\p_*\colon H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)} \to H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)};
\]
can be rewritten, after using excision to a tubular neighborhood of $\Donem$ and $\Dmone$ and
the Thom isomorphism, as a map
\[
 H_{m-1}(\Donem)\to H_{m-1}(\Dmone);
\]
the latter map is multiplication by $2$, after identifying $\Dmone$ and $\Donem$ along $p$:
indeed the normal bundle of $\Donem$ is a
double covering of the normal bundle of $\Dmone$, hence the Thom class of the first disc
bundle corresponds to twice the Thom class of the second disc bundle. We are working
with coefficients in $\Z_2$, so multiplication by $2$ is the zero map.

@@

But this is easily done by considering, in the whole argument of the proof, the surface $\N$ instead of $\mrS$:
we can define configuration spaces as in definition \ref{defn:variationsCm} using the open surface
$\N$ instead of $\mrS$, and repeat the whole argument considering this as the \emph{ambient surface}:
indeed we only need $\N$ to contain $\D$ and all curves $c_1,\dots,c_{m+1}$.

Here it is crucial that the action of $\gg$ is not involved in the statement of lemma
\ref{lem:tpsiwithproperties}, as $\N\subset\S$ is not preserved, even up to isotopy,
by diffeomorphisms of $\S$.

Lemma \ref{lem:tpsiwithproperties} is proved.

@@

Recall from the proof of lemma \ref{lem:cms->cmsDinj} that a basis for $H_m(\cms^c)\simeq H^m(\cms)$, which is
the dual of $H_m(\cms)$, is given by the classes $[0,(u_i,v_i)]$, represented by generalised
symmetric chains consisting of only one cell $e^{\tup}$, with $\tup=(0,(u_i,v_i))$.
The corresponding linear functionals are the algebraic intersection products
coming from Poincaré-Lefschetz duality:
\[
 (\cdot\cap e^{\tup})\colon H_m(\cms)\to\Z_2.
\]
Therefore
\[
 \ker\tpsi_m=\bigcap_{\tup}\ker\pa{(\cdot\cap e^{\tup})\circ\tpsi_m},
\]
and it suffices to check that $\ker\pr_m\subseteq\ker\pa{(\cdot\cap e^{\tup})\circ\tpsi_m}$
for all $\tup$ of the form $(0,(u_i,v_i))$, or equivalently, that $(\cdot\cap e^{\tup})\circ\tpsi_m$ factors through $\pr_m$.

@@

This shows that $\psi_m(c_1,\dots,c_m)$ must be $[c_1]\cdot\ldots\cdot[c_m]\in\Sym_m(\H)$, which is
one of the generating monomials.
\end{proof}

Theorem \ref{thm:Hbms*as*ggrep} is then proved in all bidegrees of the form $(m,m)$.

@@

% \begin{defn}
% %  Let $\tilde\D'$ be the open trapeziod contained in $(0,1)^2\subset\mrS$ which is delimited
% %  by the four points $(1/8;1)$, $(1/4;1/2)$, $(3/4;1/2)$ and $(7/8;1)$ (SEE PICTURE); in particular
% %  $\D\subset\D'$.
% %  Let $\S'\subset\S$ be the closure in $\S$ of $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
% %  Notice that $\S'$ is also a surface of type $\sg$.
% %  
%  Recall from section \ref{Preliminaries} that $\S'$ is the closure in $\S$ of
%  $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
%  
%  For all $m$ we define configuration spaces $C_m(\D)$ and $C_m(\S')$ as in definition \ref{defn:cms},
%  using $\D$ and $\mrS'$ instead of $\mrS$, respectively.
%  
%  For every splitting $m=p+(m-p)$ there is an open embedding
%  \[
%  \mu\colon C_p(\D)\times C_{m-p}(\S')\to \cms
%  \]
%  given by taking the union of configurations:
% 
% %  For the rest of the section let $\gg$ denote the group of isotopy classes of diffeomorphisms
% %  of $\S$ relative to $\partial\S\cup\D'$: in symbols $\gg=\pi_0\pa{\Diff(\S;\partial\S\cup\D'}$.
% \end{defn}
Recall from section \ref{Preliminaries} that $\S'$ is the closure in $\S$ of
$\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.

@@

Notice that $H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}$
is the tensor product of the trivial representation $H_{p-l}\pa{C_p(\D)}$, and
of the $\gg-$representation $H_{m-p}\pa{C_{m-p}(\S')}$, which by the results of the previous
section is isomorphic to the symplectic representation $\Sym_{m-p}(\H)$.

>>>>

\subsection{Bidigrees of the form $(m,m)$}
In this subsection we construct a $\gg-$~equivariant isomorphism 
\[
\psi_m\colon \Sym_m(\H))\to H_m(\cms).
\]
% the former group is isomorphic to $\Sym_m(H^1(\S))$ and the second to $H^m(\cms)$,
% both by Poincaré-Lefschetz duality.

We already know that these $\Z_2-$vector spaces have the same dimension:
indeed $\Sym_m(\H)$ is precisely the summand in bidegree $(m,m)$ in equation
\ref{eq:isovectorspaces}, using that a monomial
 whose
weight is \emph{equal} and not bigger than its degree cannot contain factors of the form
$Q^i\epsilon$.
% i.e. it must lie in the factor $\Sym_{\bullet}(\H)$, and if the degree
% is $m$ it must lie in $\Sym_m(\H)$.
% 
% The rough idea in constructing the map $\psi_m$ is the following: $\Sym_m(\H)$ is generated
% by elements of the form $[c_1]\cdot\ldots\cdot[c_m]$, where $c_1,\dots,c_m\colon\Sone\to\mrS$ are simple
% closed curves; we consider the product map 
% \[
% c_1\times\dots\times c_m\colon \pa{\Sone}^{\times m}\to \SP^m(\mrS)
% \]
% which carries the fundamental class of the $m-$fold torus to some class in
% $H_m(\SP^m(\mrS))$; we then lifts this class, in a nice way, to a class
% in $H_m(\cms)$.

@@

\begin{defn}
\label{defn:cCm}
Let $\cC^m$ be the the (discrete) set of isotopy classes of $m-$tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
such that
%\begin{itemize}
%  \item each $c_r\colon\Sone\to\mrS$ is smooth and transverse to each $\U_i$ and $\V_i$;
% \item
 any two curves $c_i,c_j$ intersect each other only inside $\D$.

% \item each curve $c_i$ is transverse to $\partial\D$.
%\end{itemize}
Curves are seen as maps $\Sone\to\mrS$, and the intersection of two curves is the intersection of their
images.

Two $m-$tuples of curves are isotopic if there is an ambient isotopy of $\S$ relative to $\partial\S\cup\D$
transforming one $m-$tuple into the other. In particular $\cC^m$ is more than countable.

We call \emph{multicurve} an element of $\cC^m$; by abuse of notation
$(c_1,\dots,c_m)$ will often represent its class in $\cC^m$, i.e. the corresponding multicurve.
See picture \ref{fig:defcCm}
% For each isotopy class we
% suppose to have fixed a representative $(c_1,\dots,c_m)$ made of curves that are
% also transverse to all $\U_i$'s and $\V_i$'s. We write
% $[c_1,\dots,c_m]$ for the isotopy class whose preferred representative is
% $(c_1,\dots,c_m)$.

We denote by $\ZcC{m}$ the free $\Z_2-$vector space with basis $\cC^m$.
There is a canonical surjective map $\pr_m\colon\ZcC{m}\to\Sym_m(\H)$ given by
\[
 \pr_m(c_1,\dots,c_m)=[c_1]\cdot\ldots\cdot[c_m],
\]
where $[c_i]\in\H$ is the fundamental class of the curve $c_i$.

The group $\gg$ acts both on $\ZcC{m}$, by acting on its basis $\cC^m$, and
on $\Sym_m(\H)$, symplectically. The map $\pr_m$ is $\gg-$equivariant.
\end{defn}

\begin{figure}\centering
 \includegraphics[scale=0.7]{figures/defcCm.png}
 \caption{A multicurve in $\cC^5$.}
\label{fig:defcCm}
\end{figure}

% Our aim is to define a $\gg-$equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
% factors through $\pr$ to give the desired $\gg-$equivariant map $\psi$; we will then show that
% $\psi$ is an isomorphism.

% The map $\tpsi_m$ that we construct will have two additional properties. The first
% is the following: for any
% $(c_1,\dots,c_m)$ representing a class in $\cC^m$, the homology
% class $\tpsi_m(c_1,\dots,c_m)$ can be represented by a singular cycle in $\cms$ which
% is \emph{supported} on the union $\D\cup c_1\cup \dots\cup c_m$, meaning that
% this singular cycle only hits configurations in $\cms$ whose points actually
% lie in $\D\cup c_1\cup \dots\cup c_m$.
% 
% For the second property we need a definition
\begin{defn}
 \label{defn:cmsD}
 The space $\cmsD$ is the subspace of $\SP^{m}(\mrS)$ of configurations where all points
 having multiplicity $\geq 2$ lie inside $D$.

 Recall that the $m-$fold symmetric product of an open surface is a $2m-$dimensional manifold;
 the space $\cmsD$ is an open subset
 of $\SP^m(\mrS)$, hence $\cmsD$ is also a manifold.
 See figure \ref{fig:defcmsD}.

 There is a natural, open inclusion $\iota_m\colon\cms\subset\cmsD$ and there is a natural map
  $\j_m\colon \ZcC{m}\to H_m(\cmsD)$ defined as follows: for a class $(c_1,\dots,c_m)\in\cC^m$
  the composition
  \[
   \begin{CD}
    \pa{\Sone}^{\times m} @>c_1\times\dots\times c_m >> (\mrS)^{\times m} @>>> \SP^m(\mrS)
   \end{CD}
  \]
has image in the subspace $\cmsD$; we define $\j_m(c_1,\dots,c_m)$ as the image
of the fundamental class
of the $m-$fold torus in $H_m(\cmsD)$. The result doesn't change
if we substitute $(c_1,\dots,c_m)$ with another $m-$tuple in the same multicurve.

@@

\begin{lem}
 \label{lem:cms->cmsDinj}
 The inclusion $\iota_m\colon\cms\to\cmsD$ induces an injective map
 \[
  \pa{\iota_m}_*\colon H_m(\cms)\to H_m(\cmsD).
 \]
\end{lem}
\begin{proof}
 It is equivalent to prove that the map $\iota_m^*\colon H^m(\cmsD)\to H^m(\cms)$ is surjective,
 or, using Poincaré-Lefschetz duality, that the map $\tH_m(\cmsD^c)\to \tH_m(\cms^c)$
 is surjective: this last map is induced by the map $\cmsD^c\to\cms^c$ collapsing the subspace
 $\cmsD^c\setminus\cms$ to $\infty$.

 Recall from lemma \ref{lem:gensymchain} that a $\Z_2-$basis for $\tH_m(\cms^c)$
 is given by classes $[p,(\alpha_j),(u_i,v_i)]$ with $p=0$, $\alpha_j=0$ for all $j$,
 and $\sum_{i=1}^g (u_i+v_i)=m$. Here and in the following we abbreviate $[0,(u_i,v_i)]$ for
 such classes.

@@

 It is straightforward to check that the map $\phi^{\tup}$ factors through
 $(\cmsD,\infty)$, as a map of pairs; surjectivity of $\tH_m(\cmsD^c)\to \tH_m(\cms^c)$ follows.
\end{proof}

@@

We now prove that the map $\j_m$ lifts along $(\iota_m)_*$
to a $\gg-$equivariant map $\tpsi_m$ as in the diagram.
Since $(\iota_m)_*$ is injective by lemma \ref{lem:cms->cmsDinj}, it suffices to prove that $\j_m$
lands in the image of $(\iota_m)_*$, and this last statement doesn't depend on how $\gg$ acts on these groups.

@@

\begin{defn}
\label{defn:variationsCm}
We introduce a few variations of the notion of configuration space; see figure \ref{fig:defcmsD}.
\begin{itemize} 
 \item The space $C_{1,m}(\S)$ is the subspace of $\mrS\times \cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots,P_m}}$ with $\bar P\neq P_i$ for all $i$; in other words it is
 the space of configurations of $m+1$ points, one of which is \emph{white} (meaning that
 it is distinguishable from the other points), whereas the other are \emph{black} and not distinguishable
 from each other.
%  \item The space $\fmsD$ is the subspace of $\mrS\times\cmsD$ containing all configurations
%  $(q;\set{p_1,\dots,p_m})$ where either $q\in\D$ may coincide with some $p_i$ (which can appear
%  with multipliciy $\geq 1$), or $q\not\in\D$ must be distinct from all $p_i$'s.
%  Again $q$ is thewhite point
 \item The space $\fmstwoD$ is the subspace of $\mrS\times\cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots, P_m}}$ where either $\bar P\in\D$ may coincide with \emph{exactly}
 one $P_i$, or
 $\bar P\not\in \D$ must be distinct from all $P_i$'s.
 Again $\bar P$ is called the white point.
 %We see that $C_{1,m}\subset\fmstwoD$ is an open subspace.
 \item The space $\cmstwoD$ is the subspace of $\SP^{m+1}(\mrS)$ of configurations where either all $m+1$ points
 are distinct, or there is exactly one point \emph{inside $\D$} with multiplicity $2$ and $m-1$ other points,
 somewhere in $\mrS$, with multiplicity $1$.
\end{itemize}
 \end{defn}

\begin{figure}\centering
 \includegraphics{figures/defcmsD.png}
 \caption{A configuration in each of the space introduced in definitions \ref{defn:cmsD} and \ref{defn:variationsCm}.
 Whenever a multiplicity is not specified, it is equal to 1.}
\label{fig:defcmsD}
\end{figure}

We have the following natural inclusions:
\[
C_{1,m}(\S)\subset\fmstwoD\subset\fmsD\subset\mrS\times\cmsD\subset\mrS\times\SP^m(\mrS);
\]
\[
C_{m+1}(\S)\subset\cmstwoD\subset C_{m+1}^{\D}(\S)\subset\SP^{m+1}(\mrS).
\]
All these spaces are
manifolds of dimension $2m+2$ and all inclusions are open.

@@

Fix a neighborhood $\N$ of $\D\cup c_1\cup\dots\cup c_{m+1}$.

@@

There is a natural map $\p\colon \fmstwoD\to \cmstwoD$, which converts the white point
into a black point. This map restricts to %maps $\fmstwoD\to\cmstwoD$ and 
a map $C_{1,m}(\S)\to C_{m+1}(\S)$, so that
% the latter being even a $m+1-$fold covering map;
we have a commutative diagram
 \begin{equation}\label{eq:cmstwodiagram}
  \begin{CD}
   C_{1,m}(\S) @>\subset >> \fmstwoD %@>\subset >> \fmsD
\\   @V\p VV @V\p VV %@V\p VV
\\   C_{m+1}(\S) @>\subset >> \cmstwoD %@>\subset >> C_{m+1}^D(\S)
   \end{CD}
\end{equation}

\begin{defn}
\label{defn:falsediagonals} 
Let 
\[
\Dmone=\cmstwoD\setminus C_{m+1}(\S)
\]
and similarly
\[
\Donem=\fmstwoD\setminus C_{1,m}(\S).
\]
We notice that $\p$ restricts to a homeomorphism $\Donem\to\Dmone$. Moreover both $\Dmone\subset\cmstwoD$
and $\Donem\subset\fmstwoD$ are closed submanifolds of codimension $2$, and the map $\p$ restricts to
a $2-$fold ramified covering between their respective normal bundles.
\end{defn}

@@

We first notice that there is a lift $[c_1]\otimes\fc$ to a cycle in $\pa{\fmstwoD}$,
which is defined by declaring the point in $[c_1]\cdot\fc$ that spins around $c_1$
to be white. We then notice that the right vertical map
\[
\p_*\colon H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)} \to H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}
\]
can be rewritten, after using excision to tubular neighborhoods of $\Donem$ and $\Dmone$ respectively,
and the Thom isomorphism, as a map
\[
 H_{m-1}(\Donem)\to H_{m-1}(\Dmone).
\]
The latter map is multiplication by $2$, after identifying $\Dmone$ and $\Donem$ along $p$:
indeed the normal bundle of $\Donem$ is a
double covering of the normal bundle of $\Dmone$, hence the Thom class of the first disc
bundle corresponds to twice the Thom class of the second disc bundle. We are working
with coefficients in $\Z_2$, so multiplication by $2$ is the zero map.

@@

This can be done by replacing, in the whole argument of the proof, the surface $\mrS$ with the surface $\N$.
We can define configuration spaces as in definition \ref{defn:variationsCm} also for the open surface
$\N$, and we can repeat the argument considering $\N$ as the \emph{ambient surface}:
indeed we only needed a surface containing $\D$ and all curves $c_1,\dots,c_{m+1}$.

It is crucial that the action of $\gg$ is not involved in the statement of lemma
\ref{lem:tpsiwithproperties}, as $\N\subset\S$ is not preserved, even up to isotopy,
by diffeomorphisms of $\S$.
Lemma \ref{lem:tpsiwithproperties} is proved.

@@

Recall from the proof of lemma \ref{lem:cms->cmsDinj} that a basis for $\tH_m(\cms^c)\simeq H^m(\cms)$, which is
the dual of $H_m(\cms)$, is given by the classes $[0,(u_i,v_i)]$, represented by generalised
symmetric chains consisting of only one cell $e^{\tup}$, with $\tup=(0,(u_i,v_i))$.
The corresponding linear functionals are the algebraic intersection products
coming from Poincaré-Lefschetz duality:
\[
 (\cdot\cap e^{\tup})\colon H_m(\cms)\to\Z_2.
\]
Therefore
\[
 \ker\tpsi_m=\bigcap_{\tup}\ker\pa{(\cdot\cap e^{\tup})\circ\tpsi_m},
\]
and it suffices to check that $\ker\pr_m\subseteq\ker\pa{(\cdot\cap e^{\tup})\circ\tpsi_m}$
for all $\tup$ of the form $(0,(u_i,v_i))$, or equivalently, that $(\cdot\cap e^{\tup})\circ\tpsi_m$ factors through $\pr_m$.

@@

This shows that $\psi_m(c_1,\dots,c_m)=[c_1]\cdot\ldots\cdot[c_m]\in\Sym_m(\H)$, which is
one of the generating monomials.
\end{proof}

Theorem \ref{thm:Hbms*as*ggrep} is now proved in all bidegrees of the form $(m,m)$.

@@

% \begin{defn}
% %  Let $\tilde\D'$ be the open trapeziod contained in $(0,1)^2\subset\mrS$ which is delimited
% %  by the four points $(1/8;1)$, $(1/4;1/2)$, $(3/4;1/2)$ and $(7/8;1)$ (SEE PICTURE); in particular
% %  $\D\subset\D'$.
% %  Let $\S'\subset\S$ be the closure in $\S$ of $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
% %  Notice that $\S'$ is also a surface of type $\sg$.
% %  
%  Recall from section \ref{Preliminaries} that $\S'$ is the closure in $\S$ of
%  $\S\setminus\bar\D$, where $\bar\D$ is the closure of $\D$ in $\S$.
%  
%  For all $m$ we define configuration spaces $C_m(\D)$ and $C_m(\S')$ as in definition \ref{defn:cms},
%  using $\D$ and $\mrS'$ instead of $\mrS$, respectively.
%  
%  For every splitting $m=p+(m-p)$ there is an open embedding
%  \[
%  \mu\colon C_p(\D)\times C_{m-p}(\S')\to \cms
%  \]
%  given by taking the union of configurations:
% 
% %  For the rest of the section let $\gg$ denote the group of isotopy classes of diffeomorphisms
% %  of $\S$ relative to $\partial\S\cup\D'$: in symbols $\gg=\pi_0\pa{\Diff(\S;\partial\S\cup\D'}$.
% \end{defn}
Recall from section \ref{sec:Preliminaries} that $\S'$ is the closure in $\S$ of
$\S\setminus\bar\D$.%, where $\bar\D$ is the closure of $\D$ in $\S$.

@@

Notice that $H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}$
is the tensor product of the trivial representation $H_{p-l}\pa{C_p(\D)}$, and
of the representation $H_{m-p}\pa{C_{m-p}(\S')}$, which by the results of the previous
section is isomorphic to the symplectic representation $\Sym_{m-p}(\H)$.

We will prove the following lemma, from which theorem \ref{thm:Hbms*as*ggrep} follows:
\begin{lem}
 \label{lem:oplussplitting}
For all $l\leq p\leq m$ the map $\mu_*$ in equation \ref{eq:mu*} is injective, and the collection
of all these maps yields a splitting
\begin{equation}
  \label{eq:musplitting}
 H_{m-l}(\cms)=\bigoplus_{p=l}^m H_{p-l}\pa{C_p(\D)}\otimes H_{m-p}\pa{C_{m-p}(\S')}.
\end{equation}
\end{lem}
\begin{proof}
Notice that the statement of the lemma doesn't depend on the the action of $\gg$:
we have a map from the right-hand side to the left-hand side of equation \ref{eq:musplitting},
we already know that it is $\gg-$equivariant,
we only need to show that it is a linear isomorphism. Notice also that lemma \ref{lem:gensymchain}
implies that the two vector spaces have the same dimension.
% 
% We can equivalently consider the maps in cohomology
%  \[
%    \mu^*\colon  H^{m-l}\pa{\cms}\to H^p\pa{C_{m-p}(\S')}\otimes H^{p-l}\pa{C_p(\D)},
%  \]
% and show that they are surjective for $0\leq p\leq m-l$ and together exhibit a product splitting
% \[
%   H^{m-l}(\cms)=\prod_{p=0}^{m-l} H^p\pa{C_{m-p}(\S')}\otimes H^{p-l}\pa{C_p(\D)}.
% \]
% Notice that the natural target for the cohomology map $\mu^*$ is the whole group
% \[
%  H^{m-l}\pa{C_{m-p}(\S')\times C_p(\D)}\simeq \prod_{q=0}^p H^q\pa{C_{m-p}(\S')}\otimes H^{m-l-q}\pa{C_p(\D)},
% \]
% where we have used the K\"unneth formula; we consider here, by abuse of notation,
% the composition of this map with the projection
% on the factor corresponding to $q=p$.
% 
% Using Poincaré-Lefschetz duality and again the K\"unneth formula we reduce to show that the maps
% \[
% \mu^c_* \colon \tH_{m-l}(\cms^c)\to \tH_p\pa{C_{m-p}(\S')^c}\otimes \tH_{p-l}\pa{C_p(\D)^c}
% \]
% are surjective and exhibit $\tH_{m-l}(\cms^c)$ as a product
% \[
% \tH_{m-l}(\cms^c)=\prod_{p=0}^{m-l}\tH_p\pa{C_{m-p}(\S')^c}\otimes \tH_{p-l}\pa{C_p(\D)^c}.
% \]
% Here a few explanations are required:
% we are considering the map $\mu^c\colon \cms^c\to \pa{C_{m-p}(\S')\times C_p(\D)}^c$ which collapses the complement
% of the open subspace $C_{m-p}(\S')\times C_p(\D)\subset\cms$ to the point at infinity; it induces a map
% $\mu^c_*\colon \tH_{m-l}(\cms^c) \to \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^c}$;
% applying Poincaré-Lefschetz duality
% and the K\"unneth formula to the second group we get an isomorphism
% \[
%  \tH_{m-l}\pa{\pa{C_{m-p}(\S')\times C_p(\D)}^c}\simeq \prod_{q=0}^p \tH_q\pa{C_{m-p}(\S')}\otimes\tH_{m-l-q}\pa{C_p(\D)};
% \]
% we are considering the composition of the map $\mu^c_*$ with the projection onto the factor $q=p$, and by
% abuse of notation we call this map simply $\mu^c_*$.
%

Fix $l\leq p\leq m$ and let $[a]=\prod_{j=0}^{\infty}(Q^j\epsilon)^{\alpha_j}$ be a generator of
$H_{p-l}(C_p(\D))$, hence $l=\sum_j\alpha_j$ and $p=\sum_j\alpha_j2^j$. Let also
$[b]=\prod_{i=1}^g (\u_i^{u_i}\v_i^{v_i})$ be a generator of $H_{m-p}(C_{m-p}(\S'))$, using the
isomorphism proved in the previous subsection, with $(m-p)=\sum_i(u_i+v_i)$. Here $a$ and
$b$ are chosen singular cycles representing the homology classes, with $a$ supported on $\D$
and $b$ supported on $\S'$. See figure \ref{fig:axb}
